BOLD Parameter Estimation Using Particle Filters

### 2010-07-18 doc/thesis/results.tex
more in the results

\chapter{Simulation}
Two levels are discussed in this section. First the simulation of single time series.
This is necessary to investigate the power of particle filters in 
identifying the BOLD model in a noisy environment. If the particle filter is unable
to get back to the "true" time-series in spite of noise then the particle filter
will not be useful. Additionally, determination of reasonable parameters from the
time-series is also desirable. The second simulation technique will be on a 
large group of voxels. Although the current technique treats each voxel separately,
the "Possum" FMRI simulation tool models noise more realistically, and provides additional
incite into the applicability of particle filters in modeling the BOLD signal on a large
scale.

\section{Single Timeseries}
Given the state-space equations for the BOLD signal, simulating a single time 
series is relatively straight forward. After a "true" signal is generated,
I then add identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian, I.I.D. steps. Finally a "carrier level" is added, since BOLD is typically
measured as a \% difference from the baseline. Although technically the particle filter
algorithm will immediately remove this, and calculate the \% difference, adding a carrier 
level means that the exact same algorithm which will is used for 
simulated data may be used for the real data.

Once this noisy simulated time series has been generated, the particle filter algorithm
 is then run on this single time series "image". Here I will include two sets of tests 
to determine the power of the particle filter in modeling. The first set are 
used to demonstrate the ability of the particle filter to find the most likely
set of parameters/state variables over the course of the run. The second type
of test will investigate the problem of false-positives. As the first
round of tests will show, given that there is an underlying BOLD signal, 
the particle filter is excellent at finding the most probable cause of 
the observed signal; however, given enough noise it will also determine
the most likely underlying signal; which may not be flat. Thus, it is 
important to investigate methods of determining false positives using 
average residuals.

\subsection{Signal with Low Noise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values.

To begin the single voxel simulation; I generated a signal using the following parameters:
$\{\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
These same parameters will used again in the high noise case. The noise realizations were
then based on a measurement noise, $\sigma_y$, of $.001$ and a drift standard deviation,
$\sigma_x$ of $.0005$. The measurement noise as well as the steps of the drift
were taken to be Gaussian
For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
\begin{figure}
\label{fig:LowNoiseRealization}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal}
The resulting fits compared to the \emph{true} BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
While the error between the true signal and the estimated signal are low, the noise
is also relatively low. The actual signal that was delivered into the particle filter
is the result of some preprocessing to remove drift, as described in 
\autoref{sec:Preprocessing}. The preprocessed signal is compared
to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
\begin{figure}
\label{fig:PreprocessedLowNoise}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_lownoise}
\caption{A comparison of the preprocessed signals for the low noise case. This is the
noisy input to the actual particle filter algorithm.}
The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Preprocessing}.
Obviously the spline struggles to match the signal near the end, as the graph shows, although 
overall the preprocessing seems to have been successful at removing trends. Finally, the 
set of fits to the preprocessed data, and in 
the true BOLD signal is shown in \autoref{fig:FitComparisonLowNoise}.
\begin{figure}
\label{fig:FitComparisonLowNoise}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/comparison_lownoise}
\caption{A comparison of the fitted signals for the low noise case.}
The filtering property of the particle filter is clear from the results here.  In several locations the output
of the particle filter looks like a filtered version of the input. For instance toward the
end, the estimates stay flat in spite of the preprocessed data drifting off a bit. By
this point, the algorithm had converged sufficiently to know that that such a movement in
signal isn't possible. A similar circumstance occurs at 100 seconds in. 
In general the fits are good, although because
of the slightly higher signal levels in the preprocessed signals, the fits tend to overestimate
the peaks. The final parameter sets are shown in \autoref{tab:LowNoiseParams}
\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSE}$ (Res.) &$\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.22214 & 0.3449 & 0.33462 & 0.07138 & 1.60446 & 2.27529 & 1.59454 & 5.10189  &  0.00321067  & 0.00987647  \\
1.37493 & 0.33183 & 0.36296 & 0.07327 & 1.64076 & 2.10299 & 1.57627 & 5.11867 &  0.00305475  & 0.00993227  \\
1.16604 & 0.32205 & 0.34057 & 0.08217 & 1.64768 & 2.35351 & 1.24515 & 5.16722 &  0.00328932  & 0.00967961  \\
1.23176 & 0.32707 & 0.34028 & 0.07961 & 1.62698 & 2.18519 & 1.30326 & 5.04394 &  0.00284719  & 0.00912005  \\
1.1832 & 0.31787 & 0.34718 & 0.0821 & 1.54961 & 2.29115 & 1.27817 & 5.02396   &  0.00300634  & 0.00971258  \\
1.1424 & 0.33395 & 0.34725 & 0.07366 & 1.62208 & 2.29084 & 1.4025 & 5.05531   &  0.00283287  & 0.00948483  \\
1.30041 & 0.3596 & 0.35643 & 0.07679 & 1.56406 & 2.13234 & 1.60338 & 4.99681  &  0.00302802  & 0.01021885  \\
1.24008 & 0.34601 & 0.33978 & 0.08903 & 1.64989 & 2.23655 & 1.29004 & 5.12651 &  0.00304378  & 0.01007964  \\
1.1709 & 0.32739 & 0.34644 & 0.08255 & 1.53734 & 2.28262 & 1.37828 & 4.99087  &  0.00334488  & 0.01032886  \\
1.18967 & 0.34344 & 0.33554 & 0.07976 & 1.53582 & 2.30746 & 1.42774 & 5.03295 &  0.00317542  & 0.01001503  \\
1.184 & 0.34053 & 0.35017 & 0.08917 & 1.61025 & 2.27926 & 1.16448 & 5.07352   &  0.00288855  & 0.00950536  \\
\hline                                                                           
1.21868 & 0.33588 & 0.34557 & 0.07995 & 1.59899 & 2.24884 & 1.38762 & 5.06651 & 0.00306562     & 0.00981396 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
last is mean over the 10 runs. The $\sqrt{MSE}$ (Res.) is the square root of the mean square of the
residuals. Essentially this is the is the MSE between the estimated signal and the noisy signal which 
was available to the particle filter. Square root of MSE is the actual \emph{error}, with the true signal,
which of course was not available to the particle filter.}
\label{tab:LowNoiseResults} 
\end{table}
There are a few results worth noting here. First the time constants vary greatly across
runs with different noise realizations, yet the sum of the individual time constants
($\tau_f$, $\tau_s$ and $\tau_0$) seem to be more consistent. In general the 
time constants are falling short of the true time constant. This could be a limitation
based on the prior distribution (which notably has an initial mean below the values used
in the simulation) or it could be caused by the dissassociated benefits of a correct time
constant with correct output. It often takes several measurement periods before a difference
in time constants becomes relevent. It is possible that in that time other parameters 
may be more relevent. As a result, by the time the time constant has made a substantial
difference in the weight, the 
particle with the better $\tau$ will have been weighted based on other parameters several
times. Another interesting result in the huge variation in the levels of $V_0$. In general,
with this admittedly small amount of noise, it would appear that the relation between a set
of parameters/stimuli and the output is not injective. In other words, it would appear
that a timeseries is not unique to a single set of parameters. This is good justification that 
simultaneous blood volume or tagged flow calculations with the conventional FMRI 
could definitely benefit the model. 
\begin{table}[t]
\centering
\begin{tabular}{|c | c  c  c  c  c  c  c |}
\hline 
  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
\hline 
\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
\hline 
\end{tabular}
\caption{Typical Covariance matrix of the parameters at the end of a run.}
\label{tab:CovSim} 
\end{table}

\begin{figure}[H]
\subfigure[Converging histogram for $\tau_0$, $\alpha$, $E_0$, and $V_0$ of the first run, low noise simulation.]
{\includegraphics[trim=7cm 4cm 7cm 4cm, width=15cm]{images/converge_lownoise1}}\\

\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $V$ of the first run, low noise simulation.]
{\includegraphics[trim=7cm 4cm 7cm 1cm, width=15cm]{images/converge_lownoise2}}\\
\end{figure}

\begin{figure}
\subfigure[Converging histogram for $Q$, $S$, $F$, and $BOLD$ of the first run, low noise simulation.]
{\includegraphics[trim=7cm 2cm 7cm 2cm, width=15cm]{images/converge_lownoise3}}
\label{fig:LowNoiseHist}
\end{figure}
The covariance matrix (\autoref{tab:CovSim}) also confirms the idea that the $\tau$ parameters are interchangeable
as far as the BOLD signal goes. Notice the covariance of $\tau_f$ and $\tau_0$ is $-0.019$ whereas
the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly there is 
some measure of ill-defined behavior between the differen time-constants, as well as several
other variables.  The convergence properties of the first run in \autoref{tab:LowNoiseResults} may be
viewed in \autoref{fig:LowNoiseHist}.

TODO more about convergence

\subsection{High Noise}
\label{sec:HighNoise}
For the high noise simulation, the exact same procedure was performed again, but with 
measurement error and drift standard
deviations an order of magnitude higher. The noisy signals versus the base signal are shown
in \autoref{fig:HighNoiseRealization} and the preprocessed signals which the particle
filter attempt to fit are in \autoref{fig:PreprocessedHighNoise}.
\begin{figure}[H]
\label{fig:LowNoiseRealization}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_highnoise}
\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
\end{figure}
\begin{figure}[H]
\label{fig:PreprocessedHighNoise}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\end{figure}

The results of the particle filter, for each of the ten runs may be seen in 
\autoref{fig:FitComparisonHighNoise}. Clearly the preprocessing led the algorithm
to somewhat higher activation levels, and it would appear that the subtleties of
different time constants are also lost for most of the runs, although a few seem
to manage quite accurate results.

\begin{figure}[H]
\label{fig:FitComparisonHighNoise}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/comparison_highnoise}
\caption{A comparison of the fitted signals for the high noise case.}
\end{figure}
\begin{figure}[H]
\label{fig:NoiseComparisonJustTwo}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/highnoise_56_noise}
\caption{Two particular preprocessed noise realizations for the high noise case.}
\end{figure}
\begin{figure}[H]
\label{fig:FitComparisonHighNoiseJust2}
\includegraphics[trim=6cm 3cm 6cm 3cm,width=16cm]{images/comparison_highnoise_just2}
\caption{The results for the noise realizations shown in \autoref{fig:NoiseComparisonJustTwo}.}
\end{figure}

It is interesting to consider how the preprocessing and noise may effect the
parameters of the fitting model. For instance run 1 in \autoref{fig:NoiseComparisonJustTwo}
certainly seems to be more biased toward higher peaks than run 2. There also appears
to be more drift than the 20 points per knot could fit, which explains the 
prolonged increase at 170 seconds. Despite run 1's bias toward higher peaks, for
some reason the particle filter was able to get a much better match for the
modeled post-stimulus undershoot (which is likely significantly shorter than it
should be). Ultimately the results are pretty good. \autoref{tab:SimMSE} shows
the mean squared error for all two runs, and highlights the two runs analyzed here and 
in \autoref{fig:ConvergenceRuns1} and \autoref{fig:ConvergenceRuns2}.

\begin{figure}[H]
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/highnoise_run5_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/highnoise_run5_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/highnoise_run5_3}}
\label{fig:ConvergenceRuns1}
\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
\end{figure}

There are a number of interesting convergence properties of the
particle filter when more noise is present, as both \autoref{fig:ConvergenceRuns1} and
\autoref{fig:ConvergenceRuns2} show. Obviously the particle filter seems to converge
significantly faster, as points tend to be further out on the weighting function. This
also causes significantly more resampling which is the explanation for the seeming
jumps in resolution that occur from time to time.

\begin{figure}[H]
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[trim=7cm 3cm 7cm 4cm, width=15cm]{images/highnoise_run6_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[trim=7cm 3cm 7cm 1cm, width=15cm]{images/highnoise_run6_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[trim=7cm 3cm 7cm 2cm, width=15cm]{images/highnoise_run6_3}}
\label{fig:ConvergenceRuns2}
\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
\end{figure}

The parameters arrived at for all ten filter runs are shown in \autoref{tab:HighNoiseResults}.

>>>>

\chapter{Simulation}
Two levels of simulation are discussed in this section. First the simulation of single time series
and second, a slice of simulated FMRI time series. The single time series tests
are necessary to investigate the power of particle filters in 
identifying the BOLD model in a noisy environment. If the particle filter is unable
to get back to the "true" time-series in spite of noise then the particle filter
will not be useful. Additionally, determination of reasonable parameters from the
time-series is also desirable. The reason to then simulate an entire slice is that
the "Possum" FMRI simulation tool models noise more realistically, for instance by
adding spatial correlation. Additionally 
it provides insight into the applicability of particle filters in modeling the BOLD 
signal on a large scale.

\section{Single Time Series}
Given the state-space equations for the BOLD signal, simulating a single time 
series is relatively straight forward. After a "true" signal is generated,
identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps are added to the true signal. Finally a 
"carrier level" is added, since BOLD is typically
measured as a \% difference from the baseline. The particle filter
algorithm will immediately remove this by calculating the \% difference, 
but adding a carrier level means that the exact same algorithm which 
will is used for simulated data may be used for the real data.

Once this noisy simulated time series has been generated, the particle filter algorithm
 is then run on this single time series "image". Here I will include two sets of tests 
to determine the power of the particle filter in modeling. The first set are 
used to demonstrate the ability of the particle filter to find the most likely
set of parameters/state variables over the course of the run. The second type
of test will investigate the problem of false-positives. As the first
round of tests will show, given that there is an underlying BOLD signal, 
the particle filter is excellent at finding the most probable cause of 
the observed signal. However, as discussed in the false positive tests,
even if there is no true underlying signal, the particle filter will still determine
the most likely BOLD signal; which may not be flat. Thus, it is 
important to investigate methods of determining false positives using 
average residuals. This problem could in actuality never come up, for instance
if the number of measurements is very large or if the noise peaks are not
particularly large.

\subsection{Signal with Low Noise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal.}
\label{fig:LowNoiseRealization}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_lownoise}
\caption{A comparison of the preprocessed signals for the low noise case. This is the
noisy input to the actual particle filter algorithm.}
\label{fig:PreprocessedLowNoise}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_lownoise}
\caption{A comparison of the fitted signals for the low noise case.}
\label{fig:FitComparisonLowNoise}
\end{figure}

To begin the single voxel simulation; I generated a signal using the following parameters:
$\{\tau_0 = 1.45, \alpha = .3, E_0 = .47, V_0 = .044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$.
These same parameters will be used again in the high noise case. The noise realizations were
then based on a measurement noise, $\sigma_y$, of $.001$ and a drift standard deviation,
$\sigma_x$ of $.0005$. The measurement noise as well as the steps of the drift
were taken to be Gaussian
For this low noise case, the eleven realizations are shown in \autoref{fig:LowNoiseRealization}.
The actual signal that was delivered into the particle filter
is the result of some preprocessing to remove drift, as described in 
\autoref{sec:Methods Preprocessing}. The bias introduced into the signal by preprocessing 
may have some effect on the resulting fit; thus the preprocessed signal is compared
to the true BOLD signal in \autoref{fig:PreprocessedLowNoise}.
The preprocessing consists of several steps, which are discussed in detail in \autoref{sec:Methods Preprocessing}.
Obviously the spline struggles to match the signal near the end, as the graph shows, although 
overall, the preprocessing seems to have been successful at removing trends. Finally, the 
set of fits to the true BOLD signal are shown in \autoref{fig:FitComparisonLowNoise}.
The filtering property of the particle filter is clear from the results here.  In several locations the output
of the particle filter looks like a filtered version of the input. For instance toward the
end, the estimates stay flat in spite of the preprocessed data drifting off. By
this point, the algorithm has converged sufficiently to know that that such a movement in
signal isn't possible. A similar circumstance occurs at 100 seconds in. 
In general the fits are good, although the bias introduced by pre-processing clearly
results in a slightly larger peak signal. The final parameter sets are shown in 
\autoref{tab:LowNoiseResults}.

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$ &$\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
\hline                                                                           
1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
last is mean over the 10 runs. The $\sqrt{MSR}$ is the square root of the mean square of the
residuals. Although in real data I won't distinguish between residual and error, here I will.
$\sqrt{MSE}$ is the mean squared difference between the true signal and the estimated signal.
This is the actual error, due to a combination of noise and preprocessing. }
\label{tab:LowNoiseResults} 
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{|c | c  c  c  c  c  c  c |}
\hline 
  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$ \\
\hline 
\rowcolor[gray]{.8} $\tau_0$ &   0.0189481 & -0.0014269 & -0.0011267 & -1.13e-05 & -0.0025616 & -0.0189559 & 0.0070405 \\
$\alpha$ &                       -0.0014269 & 0.0026716 & 9.93e-05 & -0.0002041 & -0.0008632 & -0.0016823 & 0.0071891 \\
\rowcolor[gray]{.8} $E_0$    &   -0.0011267 & 9.93e-05 & 0.0010701 & -0.0002277 & -0.0001177 & 0.0001013 & 0.0016972 \\
$V_0$    &                       -1.13e-05 & -0.0002041 & -0.0002277 & 0.0005401 & 4.3e-06 & 4.56e-05 & -0.0080494 \\
\rowcolor[gray]{.8} $\tau_s$ &   -0.0025616 & -0.0008632 & -0.0001177 & 4.3e-06 & 0.0128056 & 0.012878 & -0.005516 \\
$\tau_f$ &                       -0.0189559 & -0.0016823 & 0.0001013 & 4.56e-05 & 0.012878 & 0.0416927 & -0.0158182 \\
\rowcolor[gray]{.8} $\epsilon$&  0.0070405 & 0.0071891 & 0.0016972 & -0.0080494 & -0.005516 & -0.0158182 & 0.1567165 \\
\hline 
\end{tabular}
\caption{Typical Covariance matrix of the parameters at the end of a run.}
\label{tab:CovSim} 
\end{table}

\begin{figure}
\subfigure[Converging histogram for $\tau_0$, $\alpha$, $E_0$, and $V_0$ of the first run, low noise simulation.]
{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise1}}\\

\subfigure[Converging histogram for $\tau_s$, $\tau_f$, $\epsilon$, and $V$ of the first run, low noise simulation.]
{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise2}}\\
\end{figure}

\begin{figure}
\subfigure[Converging histogram for $Q$, $S$, $F$, and $BOLD$ of the first run, low noise simulation.]
{\includegraphics[clip=true,trim=7cm 3cm 6cm 3cm, width=16cm]{images/converge_lownoise3}}
\label{fig:LowNoiseHist}
\end{figure}

There are a few results worth noting here. First the time constants vary greatly across
runs with different noise realizations, yet the sum of the individual time constants
($\tau_f$, $\tau_s$ and $\tau_0$) seems to be more consistent. In general the 
time constants are falling short of the true time constant. This could be a limitation
based on the prior distribution (which notably has an initial mean below the true values) 
or it could be caused by the disassociated benefits of a correct time
constant with correct output. It often takes several measurement periods before a difference
in time constants becomes relevant to the weight. Its also possible that output is insensitive
to changes in the time constants and that varying the activation durations would 
cause the variations to be more visible. Another interesting result in the huge variation 
in the levels of $V_0$. In general,
with this admittedly small amount of noise, it would appear that the relation between a set
of parameters/stimuli and the output is not injective. In other words,  a time
series is not unique to a single set of parameters. This is good justification that 
simultaneous blood volume or tagged flow calculations with the conventional FMRI 
could benefit the model.

The covariance matrix (\autoref{tab:CovSim}) also confirms the idea that the 
$\tau$ parameters are interchangeable
as far as the BOLD signal is concerned. Notice the covariance of $\tau_f$ and $\tau_0$
is $-0.019$ whereas
the variance of $\tau_0$ and $\tau_f$ are $0.019$ and $0.04$ respectively. Clearly there is 
some measure of ill-defined behavior between the different time-constants, as well as several
other variables.  The convergence properties of the first run in \autoref{tab:LowNoiseResults} 
demonstrates the migration of parameters to match the estimates. Given the relatively high
variance of parameters, its likely that more measurements or perhaps more variation
in the stimuli could further differentiate the parameters. Its also notable that
the final mean of the parameters may not be the best point estimator, and that perhaps
using the mode would be a better way to estimate the output. Regardless, the time series'
generated from the estimated mean seems to match the correct signal very well 
(\autoref{fig:FitComparisonLowNoise}).

\subsection{High Noise}
\label{sec:HighNoise}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_highnoise}
\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:HighNoiseRealization}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\label{fig:PreprocessedHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise}
\caption{A comparison of the fitted signals for the high noise case.}
\label{fig:FitComparisonHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/highnoise_56_noise}
\caption{Two particular preprocessed noise realizations for the high noise case.}
\label{fig:NoiseComparisonJustTwo}
\end{figure}

\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise_just2}
\caption{The results for the noise realizations shown in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:FitComparisonHighNoiseJust2}
\end{figure}

For the high noise simulation, the exact same procedure was performed as with the lower
noise simulation except that $\sigma_y$ and $\sigma_x$ were set to $.01$ and $.005$,
respectively. This an order of magnitude higher than the previous tests, and indeed the
noise appears to dominate the output, as seen in \autoref{fig:HighNoiseRealization}.  
As before, the results of preprocessing are helpful in identifying the reason for
bias in the estimated output signal; thus the preprocessed signals are shown in 
\autoref{fig:PreprocessedHighNoise}. The results of the particle filter
for each of the eleven runs are shown in \autoref{fig:FitComparisonHighNoise}. 
As before, the preprocessing led the algorithm
to somewhat higher activation levels, and it would appear that the subtleties of
different time constants are being lost due to the noise. This is easily seen
by observing the post stimulus undershoots; many of the estimated time series
have no post-stimulus undershoot but rather decay directly to the base level.

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$  & $\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.1900 & 0.2349 & 0.4223 & 0.128  & 1.0147 & 2.4779 & 1.1168 & 4.6826 &0.01406 &0.01573   \\
0.9721 & 0.2190 & 0.3051 & 0.061  & 0.5780 & 1.9960 & 3.4613 & 3.5461 &0.01373 &0.01378   \\
1.5795 & 0.1415 & 0.3380 & 0.1089 & 0.5843 & 2.1247 & 1.7834 & 4.2885 &0.01275 &0.01577   \\
1.1094 & 0.2374 & 0.5349 & 0.0351 & 1.2186 & 3.0736 & 2.3504 & 5.4016 &0.01673 &0.01154    \\
1.1071 & 0.2753 & 0.3365 & 0.0316 & 1.5057 & 2.6518 & 4.1910 & 5.2646 &0.01370 &0.01222   \\
0.5803 & 0.4793 & 0.4135 & 0.1189 & 0.9756 & 3.6902 & 1.0008 & 5.2461 &0.01150 &0.01316   \\
\rowcolor[rgb]{.9,.5,.5}
1.2952 & 0.2596 & 0.2756 & 0.2595 & 1.7026 & 2.8458 & 0.6617 & 5.8436 &0.01555 &0.01790   \\
\rowcolor[rgb]{.5,.5,.9}
1.5185 & 0.2199 & 0.2835 & 0.0742 & 0.8882 & 3.0771 & 1.7393 & 5.4838 &0.01205 &0.01246   \\
0.6874 & 0.3283 & 0.3979 & 0.1561 & 1.0778 & 3.1158 & 0.6643 & 4.8810 &0.01510 &0.01258   \\
1.0170 & 0.285  & 0.3474 & 0.0567 & 1.5877 & 2.6516 & 2.2852 & 5.2563 &0.01249 &0.01343   \\
0.9925 & 0.298  & 0.3221 & 0.2094 & 0.4276 & 2.2108 & 1.0167 & 3.6308 &0.01217 &0.01506   \\
\hline                                                                          
1.0954 & 0.2708 & 0.3615 & 0.1126 & 1.0510 & 2.7196 & 1.8428 & 4.8659 &0.01362 &0.01397\\
\hline 
\end{tabular}
\caption{Estimated Parameters on 10 different runs with high noise. First row is the true parameters.
Note also that the red row is Run 1 and the blue row is Run 2, as named in  \autoref{fig:NoiseComparisonJustTwo}
and \autoref{fig:FitComparisonHighNoiseJust2}}
\label{tab:HighNoiseResults} 
\end{table}

It is interesting to consider how the preprocessing and noise may effect the
parameters of the fitting model. Consider two runs, shown in 
\autoref{fig:NoiseComparisonJustTwo}, out of the previous eleven. 
Run 1 clearly has higher peaks, in spite of the fact that both signals are based
on the same underlying signal. There also appears
to be more drift than the 20 measurements per knot could fit, which explains run 1's
prolonged increase at 170 seconds in. Interestingly, run 1 and run 2 have different
aspects of the signal that match better. Run 2 had a much better match to the 
peaks, when compared to the true signal, yet run 1 matched the post-stimulus
undershoot better. Its worth noting however, that the post-stimulus undershoots
here are much shorted than the observed prolonged post-stimulus undershoot
discussed in \autoref{sec:Post Stimulus Undershoot}. 
Despite the rather significant amount of noise, the ultimate results are 
actually rather good. \autoref{tab:HighNoiseResults} shows
the square-root mean-squared-error for all eleven runs, and highlights the two runs analyzed 
in \autoref{fig:ConvergenceRuns1} and \autoref{fig:ConvergenceRuns2}.
